Smart Representations: {R}ationality and Evolution in a Richer Environment

### 2016-06-30 paper/rev1_paperBlinded.tex
Paolo edits

\begin{abstract}
  \normalsize Standard applications of evolutionary game theory look at a single, fixed game
  and focus on the evolution of behavior for that game alone. Instead, this paper uses tools
  from evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of subjectively
  representing a decision situation, paired with a method for choosing an act based on this
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that shows how subjective preference representations in terms of regret that differ from the
  actual fitness can be evolutionarily advantageous.
\end{abstract}

@@

If agents deal with a rich and variable environment, they have to face many different choice
situations.  Standard evolutionary game models frequently simplify reality in at least two
ways. Firstly, the environment is represented as a fixed \emph{stage game}; secondly, the focus
of evolutionary selection is behavior for that stage game alone. In contrast, some argue for
studying the evolutionary competition of general \emph{choice mechanisms} in a rich and
variable environment
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche,HammStev12}. In
response to this and adding to recent like-minded approaches, this paper introduces a general
meta-game model that conservatively extends the scope of evolutionary game theory to deal with
evolutionary selection of choice mechanisms in variable environments \citep[see, inter alia,
related work
by][]{Harley1981:Learning-the-Ev,ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o,OConnor2016:Evolving-to-Gen}.\mf{please
insert more references here; maybe use a footnote right here to pick out some of the most
important related pieces and say how our contribution is novel and shiny}

A choice mechanism associates decision situations with action choices. A crucial part of a
choice mechanism is the \emph{subjective representation} of the decision situation, in
particular the manner of forming preferences and beliefs about a possibly uncertain world. To
show the usefulness of the meta-game approach, this paper asks: which preference and belief
representations are ecologically valuable and lead to high fitness? The evolution of
preferences has been subject of recent interest in theoretical economics
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. Here, we argue that questions of preference
evolution should take variability in uncertainty representation into account as well. We
demonstrate that if agents have imprecise probabilistic beliefs
\citep[e.g.][]{gardsah82,levi74,walley96}, faithful and objective representation in terms of
true, real evolutionary fitness can be outperformed by subjective (e.g., regret-based)
preference representations that deviate from the true fitness that natural selection operates
on.

@@

The paper is organized as follows. Section \ref{sec:rati--subj} sets the scene by reviewing
different perspectives no rational choice. Section~\ref{sec:subj-ut,bel,dec-rules} then
introduces the meta-game approach. In doing so it covers key notions such as choice mechanism,
decision rules and subjective representations, all with an eye towards the specific application
to the evolution of preferences in Section~\ref{sec:model}. Section~\ref{sec:results} contains
our main results for this application and Section \ref{sec:extensions} discusses some
interesting extensions.

@@

An alternative view on rationality of choice is \emph{adaptationism}
\citep[e.g.,][]{Anderson1991:Is-human-cognit,ChaterOaksford2000:The-Rational-An,HagenChater2012:Decision-Making}. Adaptationism
aims to explain rational behavior by appeal to evolutionary considerations: DMs have acquired
choice mechanisms that have proved to be adaptive with respect to the variable environment
where they have evolved. A choice mechanism can be a set of distinct heuristics (the DM's
\emph{adaptive toolbox}) that have little in common
\citep[e.g.,][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to closely relate to the literature on evolution of preferences and to the philosophical debate
about the nature of rational choice, we here suggest to think of a choice mechanism
as a map from choice situations to action choices which includes an explicit level of
\emph{subjective representation} of the situation. Specifically, a subjective representation is a general way of forming preferences and beliefs about the choice situation, and we
are most interested in the question which subjective representations, and which choice mechanisms in general, are better than others
from an evolutionary point of view.

\section{Choice Mechanisms  \& Meta-Games} 
\label{sec:subj-ut,bel,dec-rules}

We view a choice mechanism as the combination of three different things: a \textit{subjective utility} (or preference), a \textit{subjective belief} and a \textit{decision rule}. In general, the agent's action choice will depend both on the agent's utility at different possible outcomes of the choice situation and on the agent's beliefs about the realization of these outcomes. The decision rule then combines the agent's subjective utility and belief, and dictates how the agent should act in case of a choice situation whose consequences are uncertain. Therefore, a decision rule is a function that associates an action choice with the agent's utility and beliefs:
$$ \text{Decision Rule: Utility }\times\text{ Beliefs } \rightarrow \text{ Actions.}  $$
The utility of an agent can be formally expressed by a function
$u:W \times A \rightarrow \mathbb{R} $, where $A$ stands for a (finite) set of actions
available to the agent and $W$ is a (finite) set of possible states of the world. There are
many different ways to describe beliefs, but for concreteness of later applications we here
assume that the agent's beliefs are represented in terms of a (possibly singleton) set of
probability functions $\Gamma \subseteq \Delta(W) $ over the possible states of the
world. Examples of well-known decision rules from the literature that will be useful later are:
\begin{enumerate}

@@

It is worth noticing that both maxmin and maximax boil down to expected utility maximization
when the set $\Gamma$ is a singleton, and in turn expected utility maximization reduces to
Laplace rule when the belief $\mu$ is a uniform probability over the states.  \mf{do we need
  ``maximax''?}

As mentioned previously, for a choice mechanism to prescribe an action, the decision rule needs
to be given a specific subjective utility and belief as input. A \textit{subjective
  representation} of a decision problem is the pair $(u,\Gamma)$, consisting of a subjective
utility and a subjective belief. We are specifically interest in general and systematic ways of
forming such subjective representations across many different decision situations.

@@

A game is an interactive decision situation.  For a given game
$G=\langle N, (A_i,\pi^G_i)_{i \in N} \rangle$, let us denote the evolutionary payoff, or
\textit{fitness}, of player $i$ by the function
$\pi^G_i: \Pi_{i \in N} A_i \rightarrow \mathbb{R}$, where $A_i$ is player $i$'s (finite) set
of actions. For simplicity of exposition we assume that all games that are played are symmetric
two-player games where $N:=\{1,2\}$, $A_1 = A_2 $ and
$\pi^G_1(a_1, a'_2)= \pi^G_2(a'_1, a_2)=:\pi^G(a,a')$.\footnote{Since payoff functions are
  symmetric, we will simply write $\pi^G(a, a')$ for $\pi^G_1(a_1, a'_2)$ and $A:=A_1=A_2$, as
  usual. However, notice that all the definitions and results can be extended to more general
  cases.} Then, the evolutionary fitness of a subjective representation $(u_c,\Gamma_c)$ is
measured in terms of the fitness of a choice mechanism that contains it, all else equal.
Formally, the fitness payoff of a choice mechanism $c$ against choice mechanism $c'$ in a
symmetric two-player game $G=\langle \{1,2\}, A, \pi^G \rangle$ is the expected payoff of $c$
against $c'$ in $G$:

$$
F_G(c,c')= \pi^G(a_c^*(u_c^G,\Gamma_c),a_{c'}^*(u_{c'}^G,\Gamma_{c'})), \footnote{Whenever a choice mechanism
  would not select a unique action, we assume that the player chooses one of the equally
  optimal actions at random. I.e., $F_G(c,c')=
  \sum_{a \in a_c^*(u_c^G,\Gamma_c)} \sum_{a' \in a_{c'}^*(u_{c'}^G,\Gamma_{c'})}
  \frac{1}{|a_c^*(u_c^G,\Gamma_c)|} \frac{1}{|a_{c'}^*(u_{c'}^G,\Gamma_{c'})|} \pi^G(a,a')$.}
$$
where $a^*_c$ is the decision rule of choice mechanism $c$. Given the game theoretic setting,
the subjective utility $u_c$ is now a function $u_c^G:A \times A \rightarrow \mathbb{R}$, and
the subjective belief $\Gamma_c$ is a set of probability functions over the co-player's
actions, $ \Gamma \subseteq \Delta(A)$.

However, studying the fitness and the evolutionary selection of different choice mechanisms by
the use of a fixed, single game may look like a rather restrictive perspective and possibly
biased by the choice of the particular game. To avoid these limitations, we consider a
\textit{class} of possible games that can be played within a population
(REFERENCES). \mf{please insert some references to other papers that use a multi-game approach
  as well.} Towards this end, fix a class $\mathcal{G}$ of possible two-player symmetric games,
together with a probability measure $P_{\mathcal{G}}(G)$ for the occurrence probability of game
$G \in \mathcal{G}$. The probability measure $P_{\mathcal{G}}$ encodes the statistical
properties of the environment. A \textit{meta-game} is the tuple
$ MG=\langle CM, \mathcal{G}, P_{\mathcal{G}},F \rangle$, where $CM$ is a set of choice
mechanisms, $\mathcal{G}$ is a class of possible games, $P_{\mathcal{G}}(G)$ is the probability
of game $G$ to occur, and $F:CM \times CM \rightarrow \mathbb{R}$ is the fitness function,
defined as:
\begin{align}
  \label{eq:FittnessChoiceMechGamePairwise}
  F(c, c') = \int P_{\mathcal{G}}(G) \  F_G(c,c') \text{ d} G \,.
\end{align}
Hence, $F(c,c')$ determines the evolutionary payoff of choice mechanism $c$ against $c'$ in the
meta-game. The set $CM$ can be thought of as the set of choice mechanisms present within a given
population that is playing the games from the class $\mathcal{G}$. Consequently, it is possible
to compute the average fitness of $c$ against the population, that is given by:
\begin{align}
  \label{eq:FittnessChoiceMechGame}
  F(c) = \int P_{M}(c') \ F(c,c') \text{ d} c' = \int \int P_{\mathcal{G}}(G) \  P_{M}(c') \  F_G(c,c') \text{ d} c' \text{ d} G  \,,
\end{align}
where $P_{M}(c')$ is the probability of
encountering a co-player with choice mechanism $c'$.

Meta-games are then abstract models for the evolutionary competition between choice mechanisms
in interactive decision making contexts.  Standard notions of evolutionary game theory apply to
meta-games as well. For example, a choice mechanism $c$ is a strict Nash equilibrium if
$F(c,c) > F(c',c)$ for all $c'$; it is evolutionarily stable if for all $c'$ either (i)
$F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$ and $F(c,c') > F(c',c')$; it is neutrally stable
if for all $c'$ either (i) $F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$ and
$F(c,c') \ge F(c',c')$ \citep{Maynard-Smith1982:Evolution-and-t}. Similarly, evolutionary
dynamics can be applied to meta-games. We will later turn towards the replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St} and the replicator mutator dynamic
\citep[e.g.][]{Nowak2006:Evolutionary-Dy}.

\section{Example Application: Evolution of Preferences} \label{sec:model}

To demonstrate the potential usefulness of a meta-game approach, we will compare a small
selection of general ways of forming utility and preference representations against each
other. As for utilities, consider initially:
\begin{enumerate}
\item the \textit{objective} utility, defined by: for all $G \in \mathcal{G}$,
$$\text{obj}^G(a,a')=\pi^G(a,a');$$
\item the \textit{regret}, defined by: for all $G \in \mathcal{G}$,
$$\text{reg}^G(a,a') =\pi^G(a,a')- \max_{a''\in A} \pi^G(a'',a').$$
\end{enumerate}

For a start, the subjective beliefs that we take into consideration are also two:
\begin{enumerate}
\item $prc$: a precise uniform belief $\overline{\mu}$ such that $\overline{\mu}(a)=\frac{1}{|A|}$ for all $a\in A$;
\item $imp$: a maximally imprecise belief $\overline{\Gamma}=\Delta(A)$.
\end{enumerate}

Although a thorough discussion of this issue goes beyond the scope of this work, let us say
that these two kinds of belief underlie two different and alternative views on
uncertainty. Faced with uncertain events, a strict Bayesian will always form a precise belief,
specified by a \textit{single} probability $\mu$. In the absence of any information about
future uncertain events, the Bayesian would then invoke the \textit{principle of insufficient
  reason}, and accordingly choose a uniform probability over the possible outcomes. In
contrast, others have argued against the obligation of representing a belief by means of a
single probability measure, opposite to the Bayesian prescriptions (REFERENCES). They argue
instead in favor of a more general account, according to which uncertainty can be unmeasurable,
and represented by a (closed and convex) set of probabilities. This line of thought has been
originated in decision theory by the famous Ellsberg's paradoxes (REF), and appears extremely
relevant in game-theoretical contexts too. Indeed, in a recent paper \citet{BattCerrMM15}
write:

\noindent In evolutionary game theory, for instance, players will obviously face uncertainty
about the composition of the population that they are part of, and consequently about the
co-player that they are randomly paired with at each round and about the co-player's action. In
case of complete lack of information, a non-Bayesian player would thus entertain maximal
unmeasurable uncertainty, i.e., a maximally imprecise belief.\footnote{Such a radical
  uncertainty could ensue, for example, if agents had no conception of their co-player or her
  preferences. Unsophisticated agents, as considered in evolutionary game theory, might be
  entirely unaware of the fact that they are engaged in social decision making \cite[e.g.,][for
  game theoretic models of unawareness]{HeifetzMeier2009:Dynamic-Unaware}. IT is therefore
  certainly not entirely ludicrous to consider radical uncertainty first and tend to more
  sophisticated ways of forming beliefs later (more on this below).} As already anticipated, we
will see that the way agents can form beliefs, and the possibility of holding imprecise beliefs
in particular, may have a fundamental impact on their evolutionary success.

As for the decision rule, we assume that the players use the maxmin rule. This is in line with most of the representation
results of decision making under uncertainty \citep[e.g.,][]{gilsch89,KlibMarMuk05,GhirMar02},
and seems justified by empirical observations too. Ellsberg's paradoxes are prominent examples (\citet{ells61}), and evidence from experimental literature suggests that agents are mostly uncertainty averse \citep[e.g.,][]{TrautKuil16}.

Finally, note that when the maxmin rule acts on subjective representation (obj, imp), the
generated behavior corresponds to the classic maxmin (REF:vNM). When the maxmin rule acts on
subjective representation (reg, imp), the agent's behavior is known as \textit{regret
  minimization}.\footnote{The notion of regret in decision theory dates back at least to the
  work by \citet{Savage1951:The-theory-of-s}, and has later been developed by \citet{bell82},
  \citet{fish82} and \citet{loosug82} independently. Recently,
  \citet{HalpernPass2012:Iterated-Regret} showed how the use of regret minimization can give
  solutions to game theoretical-puzzles (like the Traveller's dilemma and the Centipede game)
  in a way that is closer to everyday intuition and empirical data. In this paper the notion of
  regret defined earlier is the same as in \citet{HalpernPass2012:Iterated-Regret}.} Two facts
stem from these observations.  The first is related to our focus on different types of
uncertainty that players may entertain.
\begin{fact} \label{fact:maxEU-minReg} 
%Under an identical precise (singleton set) belief, the
%  acts selected by maximization of expected evolutionary payoff in game $G$ are exactly the acts selected
%  by the maximization of expected utility for the negative regret transformation of $G$
Under any identical precise (Bayesian) belief, maximization of expected evolutionary payoff and minimization of expected regret are behaviorally equivalent. %\citep[e.g.,][]{HalpernPass2012:Iterated-Regret}.
\end{fact}
\noindent The second fact highlights another behavioral equivalence, which we will make use of shortly in the following section.
\begin{fact} \label{fact:equivalence2x2} In the class of $2 \times 2$ symmetric games, the acts
  selected by Laplace rule are exactly the acts selected by regret minimization.
\end{fact}

Here is a simple example that shows these choice mechanisms in action. Consider the
coordination game in evolutionary payoffs $G$ depicted in figure \ref{coordgame1}. Since the
game is symmetric, it suffices to specify the payoffs for the row player. Figure
\ref{coordgame1} also represents the objective utility $\text{obj}^G$, since
$\text{obj}^G= \pi^G$ by definition, whereas figure \ref{coordgame1reg} pictures the
representation of $G$ in terms of regret-based utilities. While classic maxmin is indifferent
between $I$ and $II$ (figure \ref{coordgame1}), regret minimization uniquely selects $II$
(figure \ref{coordgame1reg}).
\begin{figure}
\begin{center}
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tabular}{ccc}
      \toprule
      & I & II \\
      \midrule
      I & 1 & 0 \\
      II & 0 & 2\\
      \bottomrule
    \end{tabular}
    \caption{Coordination game $G$.}
    \label{coordgame1}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \begin{tabular}{ccc}
      \toprule
      & I & II \\
      \midrule
      I & 0 & -2 \\
      II & -1 & 0\\
      \bottomrule
    \end{tabular}
    \caption{Regret-based representation of $G$.}
    \label{coordgame1reg}
  \end{subfigure}
  \caption{A coordination game (left) and the associated regret representation (right).}
    \label{coordgame1mainFig}
    \end{center}
\end{figure}

\section{Some Results for the Example Scenario}\label{sec:results}

>>>>

\begin{abstract}
  \normalsize Standard applications of evolutionary game theory look at a single, fixed game
  and focus on the evolution of behavior for that game alone. Instead, this paper uses tools
  from evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable multi-game environment. A choice mechanism is a way of subjectively
  representing a decision situation, paired with a method for choosing an act based on this
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that shows how subjective representations in terms of regret that differ from the
  actual fitness can be evolutionarily advantageous.
\end{abstract}

@@

If agents deal with a rich and variable environment, they have to face many different choice
situations.  Standard evolutionary game models frequently simplify reality in at least two
ways. Firstly, the environment is represented as a fixed \emph{stage game}; secondly, the focus
of evolutionary selection is behavior for that stage game alone. In contrast, some argue for
studying the evolutionary competition of general \emph{choice mechanisms} in a rich and
variable environment
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche,HammStev12}. In
response to this and adding to recent like-minded approaches, this paper introduces a general
\emph{meta-game} model that conservatively extends the scope of evolutionary game theory to deal with
evolutionary selection of choice mechanisms in variable environments \citep[see also][]{Harley1981:Learning-the-Ev,ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o,OConnor2016:Evolving-to-Gen,RayoBecker07,Zollman2008,SkyrmsZollman10,BednarPage07}.\footnote{In particular, similar ideas appear in \citep{RayoBecker07,Zollman2008,BednarPage07}. \citep{BednarPage07} use a multi-game framework, composed of a fixed selection of six possible games, to study the emergence of different cultural behaviors, and model agents as finite-state automata playing games from the fixed selection. \citep{Zollman2008} explains seemingly ``irrational'' fair behavior in social dilemmas (like the Ultimatum game) by means of a model where agents have to play the Ultimatum game together with the Nash bargaining game, but they are constrained to choose the same strategy for both games. Finally, \citep{RayoBecker07} consider, in a more decision-theoretic setting, what subjective utility function a cognitively limited agent should be endowed with in order to maximize her evolutionary fitness. Our framework can then be viewed as a generalization of those models, mainly in that here players do not suffer any cognitive constraint, and we allow for larger and possibly variable classes of games.}

A choice mechanism associates decision situations with action choices. A crucial part of a
choice mechanism is the \emph{subjective representation} of the decision situation, in
particular the manner of forming preferences and beliefs about a possibly uncertain world. To
show the usefulness of the meta-game approach, this paper asks: which preference and belief
representations are ecologically valuable and lead to high fitness? The evolution of
preferences has been subject of recent interest in theoretical economics
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. Here, we argue that questions of preference
evolution should take variability in uncertainty representation into account as well. We
demonstrate that if agents have \emph{imprecise} probabilistic beliefs
\citep[e.g.,][]{gardsah82,levi74,walley96}, faithful and objective representations in terms of
true, real evolutionary fitness can be outperformed by subjective (e.g., regret-based)
preference representations that deviate from the true fitness that natural selection operates
on.

@@

The paper is organized as follows. Section \ref{sec:rati--subj} sets the scene by reviewing
different perspectives on rational choice. Section~\ref{sec:subj-ut,bel,dec-rules}
introduces the meta-game approach. In doing so, it covers key notions such as choice mechanisms,
decision rules and subjective representations, all with an eye towards the evolutionary applications
of Section~\ref{sec:model}. Section~\ref{sec:results} contains
the main results, and Section \ref{sec:extensions} discusses some
interesting extensions. Finally, Section~\ref{sec:conclusion} concludes.

@@

An alternative view on rationality of choice is \emph{adaptationism}
\citep[e.g.,][]{Anderson1991:Is-human-cognit,ChaterOaksford2000:The-Rational-An,HagenChater2012:Decision-Making}. Adaptationism
aims to explain rational behavior by appealing to evolutionary considerations: DMs have acquired
choice mechanisms that have proved to be adaptive with respect to the variable environment
where they have evolved. A choice mechanism can be a set of distinct heuristics (the DM's
\emph{adaptive toolbox}) that have little in common
\citep[e.g.,][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to closely relate to the literature on evolution of preferences and to the philosophical debate
about the nature of rational choice, we here suggest to think of a choice mechanism
as a map from choice situations to action choices which includes an explicit level of subjective representation of the situation. Specifically, a \emph{subjective representation} is a general way of forming preferences and beliefs about the choice situation, and we
are most interested in the question which subjective representations, and which choice mechanisms in general, are better than others
from an evolutionary point of view.

\section{Choice Mechanisms and Meta-Games} 
\label{sec:subj-ut,bel,dec-rules}

We view a choice mechanism as the combination of three different things: a \textit{subjective utility} (or preference), a \textit{subjective belief} and a \textit{decision rule}. In general, the agent's action choice will depend both on the agent's utility at different possible outcomes of the choice situation and on the agent's beliefs about the realization of these outcomes. The decision rule then combines the agent's subjective utility and belief, and dictates how the agent should act in case of a choice situation whose consequences are uncertain. Therefore, a decision rule is a function that associates an action choice with the agent's utility and beliefs:
$$ \text{Decision Rule: Utility }\times\text{ Beliefs } \rightarrow \text{ Actions.}  $$
The subjective utility of an agent can be formally expressed by a function
$u:W \times A \rightarrow \mathbb{R} $, where $A$ stands for a (finite) set of actions
available to the agent and $W$ is a (finite) set of possible states of the world. There are
many different ways to describe beliefs, but for concreteness of later applications we here
assume that the agent's beliefs are represented in terms of a (possibly singleton) convex compact set of
probability functions $\Gamma \subseteq \Delta(W) $ over the possible states of the
world. Given a utility $u$ and a belief $\Gamma$, examples of well-known decision rules from the literature that we will encounter later are:
\begin{enumerate}

@@

It is worth noticing that both maxmin and maximax boil down to expected utility maximization
when the set $\Gamma$ is a singleton, and in turn expected utility maximization reduces to
Laplace rule when the belief $\mu$ is a uniform probability over the states.

As mentioned previously, for a choice mechanism to prescribe an action, the decision rule needs
to be given a specific utility $u$ and belief $\Gamma$ as input. We call the pair $(u,\Gamma)$ a \textit{subjective
  representation} of the decision problem. In the following, we investigate the evolutionary fitness of general and systematic ways of
forming such subjective representations across many different decision situations.

@@

A \emph{fitness game} is an interactive decision situation.  For a given fitness game
$G=\langle N, (A_i,\pi^G_i)_{i \in N} \rangle$, let us denote the evolutionary payoff, or
\textit{fitness}, of player $i$ by the function
$\pi^G_i: \Pi_{i \in N} A_i \rightarrow \mathbb{R}$, where $A_i$ is player $i$'s (finite) set
of actions. For simplicity of exposition we assume that all games that are played are symmetric
two-player games where $N:=\{1,2\}$, $A_1 = A_2 $ and
$\pi^G_1(a_1, a'_2)= \pi^G_2(a'_1, a_2)=:\pi^G(a,a')$.\footnote{Since payoff functions are
  symmetric, we simply write $\pi^G(a, a')$ for $\pi^G_1(a_1, a'_2)$ and $A:=A_1=A_2$, as
  usual. However, notice that all definitions and results can be extended to more general
  cases.} The fitness of a choice mechanism $c$ with decision rule $a_{c}^{*}$ and subjective representation $(u_c,\Gamma_c)$ is
measured in terms of the expected evolutionary payoff of $c$.
Formally, the fitness of choice mechanism $c$ against choice mechanism $c'$ in a
symmetric two-player game $G=\langle \{1,2\}, A, \pi^G \rangle$ is given by:

$$
F_G(c,c')= \pi^G(a_c^*(u_c^G,\Gamma_c),a_{c'}^*(u_{c'}^G,\Gamma_{c'})). \footnote{Whenever a choice mechanism
  would not select a unique action, we assume that the player chooses one of the equally
  optimal actions at random. I.e., $F_G(c,c')=
  \sum_{a \in a_c^*(u_c^G,\Gamma_c)} \sum_{a' \in a_{c'}^*(u_{c'}^G,\Gamma_{c'})}
  \frac{1}{|a_c^*(u_c^G,\Gamma_c)|} \frac{1}{|a_{c'}^*(u_{c'}^G,\Gamma_{c'})|} \pi^G(a,a')$.}
$$
Given the game theoretic setting,
the subjective utility $u^{G}_c$ is now a function $u_c^G:A \times A \rightarrow \mathbb{R}$, and
the subjective belief $\Gamma_c$ is a set of probability functions over the co-player's
actions, $ \Gamma_c \subseteq \Delta(A)$.

However, studying the fitness and the evolutionary selection of different choice mechanisms by
the use of a fixed, single game may look like a rather restrictive perspective and possibly
biased by the choice of the particular game. To avoid these limitations, we consider a
\textit{class} of possible games. Towards this end, fix a class $\mathcal{G}$ of possible two-player symmetric games,
together with a probability measure $P_{\mathcal{G}}(G)$ for the occurrence probability of game
$G \in \mathcal{G}$. Intuitively, the probability $P_{\mathcal{G}}$ encodes the statistical
properties of the environment. A \textit{meta-game} is then a tuple
$ MG=\langle CM, \mathcal{G}, P_{\mathcal{G}},F \rangle$, where $CM$ is a set of choice
mechanisms, $\mathcal{G}$ is a class of possible games, $P_{\mathcal{G}}(G)$ is the probability
of game $G$ to occur, and $F:CM \times CM \rightarrow \mathbb{R}$ is the (meta-)fitness function,
defined as:
\begin{align}
  \label{eq:FittnessChoiceMechGamePairwise}
  F(c, c') = \int P_{\mathcal{G}}(G) \  F_G(c,c') \text{ d} G \,.
\end{align}
Hence, $F(c,c')$ determines the evolutionary payoff of choice mechanism $c$ against $c'$ in the
meta-game. The set $CM$ can be thought of as the set of choice mechanisms that are present within a given
population playing the games from the class $\mathcal{G}$. Consequently, it is possible
to compute the average fitness of $c$ against the population, that is given by:
\begin{align}
  \label{eq:FittnessChoiceMechGame}
  F(c) = \int P_{M}(c') \ F(c,c') \text{ d} c' = \int \int P_{\mathcal{G}}(G) \  P_{M}(c') \  F_G(c,c') \text{ d} c' \text{ d} G  \,,
\end{align}
where $P_{M}(c')$ is the probability of
encountering a co-player with choice mechanism $c'$.

Meta-games are then abstract models for the evolutionary competition between choice mechanisms
in interactive decision making contexts.  Standard notions of evolutionary game theory apply to
meta-games as well. For example, a choice mechanism $c$ is a strict Nash equilibrium if
$F(c,c) > F(c',c)$ for all $c'$; it is evolutionarily stable if for all $c'$ either (i)
$F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$ and $F(c,c') > F(c',c')$; it is neutrally stable
if for all $c'$ either (i) $F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$ and
$F(c,c') \ge F(c',c')$ \citep{Maynard-Smith1982:Evolution-and-t}. Similarly, evolutionary
dynamics can be applied to meta-games. Later we will also turn towards a dynamical analysis in terms of replicator dynamics
\citep{TaylorJonker1978:Evolutionary-St} and replicator mutator dynamics
\citep[e.g.,][]{Nowak2006:Evolutionary-Dy}.

\section{Evolution of Preferences} \label{sec:model}

To demonstrate the usefulness of a meta-game approach, we compare a
selection of general ways of forming belief and preference representations against each
other. As for subjective preferences, consider initially:
\begin{enumerate}
\item the \textit{objective} utility, defined by: for all $G \in \mathcal{G}$,
$$\text{obj}^G(a,a')=\pi^G(a,a');$$
\item the \textit{regret}, defined by: for all $G \in \mathcal{G}$,
$$\text{reg}^G(a,a') =\pi^G(a,a')- \max_{a''\in A} \pi^G(a'',a').$$
\end{enumerate}

As motivation for this comparison, it is to be stressed that regret minimization is one of the main alternatives to utility (or value) maximization in the literature on decision criteria (see also \cite{bleWakker15}).
For a start, the subjective beliefs that we take into consideration are also two:
\begin{enumerate}
\item $prc$: a precise uniform belief $\overline{\mu}$ such that $\overline{\mu}(a)=\frac{1}{|A|}$ for all $a\in A$;
\item $imp$: a maximally imprecise belief $\overline{\Gamma}=\Delta(A)$.
\end{enumerate}

Although a thorough discussion of this issue goes beyond the scope of this work, let us say
that these two kinds of belief underlie two different and alternative views on
uncertainty. Faced with uncertain events, a strict Bayesian will always form a precise belief,
specified by a \textit{single} probability $\mu$. In the absence of any information about
future uncertain events, the Bayesian would mostly invoke the \textit{principle of insufficient
  reason}, and accordingly choose a uniform probability over the possible outcomes. In
contrast, others have argued against the obligation of representing a belief by means of a
single probability measure, opposite to the Bayesian paradigm (e.g., \cite{GilbMarin13}). They argue
instead in favor of a more encompassing account, according to which uncertainty can be \emph{unmeasurable},
and represented by a (convex and compact) set of probabilities (e.g., \cite{gilsch89}). This line of thought has been
originated in decision theory by the famous Ellsberg's paradoxes \citep{ells61}, and appears extremely
relevant in game-theoretic contexts too. Indeed, in a recent paper \citet{BattCerrMM15}
write:

\begin{quote}
  Such [unmeasurable] uncertainty is inherent in situations of strategic interaction. This is quite obvious
  when such situations have been faced only a few times. (p.~646)
\end{quote}

\noindent In evolutionary game theory, for instance, players obviously face uncertainty
about the composition of the population that they are part of, and consequently about the (type of)
co-player that they are randomly paired with at each round and about the co-player's action. In
case of complete lack of information about the composition of the population, a non-Bayesian player would thus entertain maximal
unmeasurable uncertainty, i.e., a maximally imprecise belief.\footnote{Such a radical
  uncertainty could ensue, for example, if agents have no conception of their co-player or her
  preferences. Unsophisticated agents, as considered in evolutionary game theory, might be
  entirely unaware of the fact that they are engaged in social decision making (see \cite{HeifetzMeier2009:Dynamic-Unaware}, for
  game-theoretic models of unawareness). It is therefore
  not ludicrous to consider radical uncertainty first and tend to more
  sophisticated ways of forming beliefs later (more on this below).} As already anticipated, we
will see that the way agents form beliefs, and the possibility of holding imprecise beliefs
in particular, can have a fundamental impact on their evolutionary success.

As for the decision rule, we assume that players use the maxmin rule. This is in line with many representation
results of decision making under unmeasurable uncertainty \citep[e.g.,][]{gilsch89,GhirMar02},
and seems corroborated by empirical findings too. Ellsberg's paradoxes are prominent examples (\citet{ells61}), and evidence from experimental literature suggests that agents are generally averse to unmeasurable uncertainty \citep[e.g.,][]{TrautKuil16}.

Finally, note that when the maxmin rule acts on subjective representation (obj, imp), the
generated behavior corresponds to the classic maxmin strategy \citep{neumorg44}. When the maxmin rule acts on
subjective representation (reg, imp), the agent's behavior is known as \textit{regret
  minimization}.\footnote{The notion of regret in decision theory dates back at least to the
  work by \citet{Savage1951:The-theory-of-s}, and has later been developed by \citet{bell82},
  \citet{Fishburn1982} and \citet{loosug82} independently. Recently,
  \citet{HalpernPass2012:Iterated-Regret} showed how the use of regret minimization can give
  solutions to game-theoretic puzzles (like the Traveller's dilemma and the Centipede game)
  in a way that is closer to everyday intuition and empirical data. In this paper the notion of
  regret defined earlier is the same as in \citet{HalpernPass2012:Iterated-Regret}.} Two facts
stem from these observations.  The first is related to our focus on different types of
uncertainty that players may entertain.
\begin{fact} \label{fact:maxEU-minReg} 
%Under an identical precise (singleton set) belief, the
%  acts selected by maximization of expected evolutionary payoff in game $G$ are exactly the acts selected
%  by the maximization of expected utility for the negative regret transformation of $G$
For any precise (Bayesian) and identical belief, maximization of expected (objective) utility and minimization of expected regret are behaviorally equivalent. %\citep[e.g.,][]{HalpernPass2012:Iterated-Regret}.
\end{fact}
\noindent The second fact highlights another behavioral equivalence, which we will make use of shortly in the following section.
\begin{fact} \label{fact:equivalence2x2} In the class of $2 \times 2$ symmetric games, the acts
  selected by Laplace rule are exactly the acts selected by regret minimization.
\end{fact}

Here is a simple example that shows these choice mechanisms in action. Consider the
coordination fitness game $G$ depicted in figure \ref{coordgame1}. Since the
game is symmetric, it suffices to specify the evolutionary payoffs for the row player. Figure
\ref{coordgame1} also represents the objective utility $\text{obj}^G$, since
$\text{obj}^G= \pi^G$ by definition, whereas figure \ref{coordgame1reg} pictures the
representation of $G$ in terms of regret-based utilities. While classic maxmin is indifferent
between $I$ and $II$ (figure \ref{coordgame1}), regret minimization uniquely selects $II$
(figure \ref{coordgame1reg}).
\begin{figure}
\begin{center}
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tabular}{ccc}
      \toprule
      & I & II \\
      \midrule
      I & 1 & 0 \\
      II & 0 & 2\\
      \bottomrule
    \end{tabular}
    \caption{Coordination game $G$.}
    \label{coordgame1}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \begin{tabular}{ccc}
      \toprule
      & I & II \\
      \midrule
      I & 0 & -2 \\
      II & -1 & 0\\
      \bottomrule
    \end{tabular}
    \caption{Regret-based representation of $G$.}
    \label{coordgame1reg}
  \end{subfigure}
  \caption{A coordination game (left) and the associated regret representation (right).}
    \label{coordgame1mainFig}
    \end{center}
\end{figure}

\section{Results}\label{sec:results}
